Analysis-ready VCF at Biobank scale using Zarr

### 2025-02-06 paper.tex
Add SC2 Figshare reference

\subsection{Case study: SARS-CoV-2 alignment data}
The SARS-CoV-2 pandemic resulted in viral sequence data collection at
an unprecedented scale, with over 15 million whole genomes incorporated into
one resource as of 2024~\citep{hinrichs2024ongoing}. Data at such scale
overwhelmed classical phylogenetic methods, leading to the development
of a new class of tools~\citep{turakhia2021ultrafast,ye2022matoptimize,zhan2023towards}
capable of handling millions of samples.
The Viridian dataset~\citep{hunt2024addressing} is a consistently assembled
set of consensus sequences for 4,484,157 samples, addressing a range of systematic
errors. The \texttt{sc2ts} inference tool uses VCF-Zarr as its
input format, and Zhan et al have made the Viridian dataset,
aligned to the reference using MAFFT~\citep{katoh2013mafft},
available for download~\citep{zhan2023towards}. This dataset has quite
different properties to those examined in previous case studies, and
illustrates the flexibility of Zarr across a range of biological applications.

>>>>

\subsection{Case study: SARS-CoV-2 alignment data}
The SARS-CoV-2 pandemic resulted in viral sequence data collection at
an unprecedented scale, with over 15 million whole genomes incorporated into
one resource as of 2024~\citep{hinrichs2024ongoing}. Data at such scale
overwhelmed classical phylogenetic methods, leading to the development
of a new class of tools~\citep{turakhia2021ultrafast,ye2022matoptimize,zhan2023towards}
capable of handling millions of samples.
The Viridian dataset~\citep{hunt2024addressing} is a consistently assembled
set of consensus sequences for 4,484,157 samples, addressing a range of systematic
errors. The \texttt{sc2ts} inference tool uses VCF-Zarr as its
input format, and Zhan et al~\citep{zhan2023towards} have made the Viridian dataset,
aligned to the reference using MAFFT~\citep{katoh2013mafft},
available for download~\citep{kelleher2025viridian}. This dataset has quite
different properties to those examined in previous case studies, and
illustrates the flexibility of Zarr across a range of biological applications.
